---
title: "Dimentionality_reduction & Feature_selection"
author: "Mildred Kulei"
date: "9/10/2021"
output:
  pdf_document: default
  html_document: default
  
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tinytex)
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```


# 1. Problem Definition
## 1.1 Specifying the Question
### what is the most relevant marketing strategies that will result in the highest no. of sales (total price including tax) in carrefour Kenya.

## 1.2 Metric for success
### Come up with an analysis that will make our client identify the relevant marketing strategies to increase the company's sales.

## 1.3 Understanding the Context
### Carrefour was launched in kenya in 1999 and today it operates five hypermarkets and five supermarkets, employing over 1800 colleagues.

###Carrefour operates different store formats, to meet the growing needs of its diversified customer base. In line with the brand’s commitment to provide the widest range of quality products and value for money, Carrefour offers an unrivalled choice of more than 30,000 food and non-food products, and an exemplary customer experience to create great moments for everyone every day. In respect to this, the marketing team wants to be informed on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax). 



## 1.4 Experimental Design taken
1. Problem Definition
2. Data Sourcing
3. Check the Data
4. Perform Data Cleaning
5. Perform Exploratory Data Analysis  
6. Implement the Solution
7. Challenge the Solution
8. Follow up Questions

## 1.5 Data relevance 
### The data collected is relevant as it is was provided by the client from the company's database. 

# 2. Data Sourcing
## Loading the data
## Loading the necessary packages
``` {r}
library("data.table")
sales1 <- read.csv("Supermarket_Dataset_1 - Sales Data.csv")
```

# 3. checking the data

```{R}
##Previewing the first 6 rows of dataset

head(sales1)

##Previewing the last 6 rows of dataset

tail(sales1)

##Basic structure of the data
str(sales1)
```

```{r}
# previewing the column names
colnames(sales1)
```

```{r}
# previewing the dataset
class(sales1)
```

```{r}
# previewing the datatypes of the dataset
sapply(sales1, class)
```

```{r}
# checking the shape of the data
dim(sales1)
```
The dataset has 1000 records of data and 16 columns.

# 4. Perform Data Cleaning

## missing values
```{R}
# checking for missing values
sum(is.na(sales1))
```
no missing values.

## Duplicates
```{r}
# Identifying duplicates
duplicates <- sales1[duplicated(sales1), ]
head(duplicates)
```
There are no duplicates

## Outliers

```{r}
#install.packages("dplyr")    # alternative installation of the %>%

library(dplyr)
```


```{R}
numeric_df <- sales1 %>% select_if(is.numeric)
boxplot(numeric_df)
```
We notice that the column Cogs and Total have a few outliers

## changing column names to make them uniform
```{r}
colnames(sales1) = tolower(colnames(sales1))
```

```{r}
#to confirm the change
names (sales1)
```
## Dropping columns that we do not need

```{r}
#gross.income and tax are the same so we will drop one, 
#we also don't need invoice id in our analysis.

sales1[ ,c('tax','invoice.id')] <- list(NULL)
```

```{r}
#to confirm they have been deleted
names(sales1)
```

```{r}
#separating the date column into 'year','month','day'
library(tidyr)

sales1_date <- separate(sales1, date, c("month", "day", "year"))
head(sales1_date,n=2) 
```


We notice that our data is from the year 2019

```{r}
library(tidyverse)
library(magrittr) 

#Factors are used to represent categorical data and are important for statistical analysis and for plotting.
#For the categorical data we will change to levels

sales1_cat= c('branch', 'customer.type', 'gender', 'product.line', 'payment', 'year','month','day')

# Changing columns to factors

sales1_date[,sales1_cat] %<>% lapply(function(x) as.factor(as.character(x)))
str(sales1_date)
```
We observe that:
there rae two types of customer; member and normal.
There are three branches.
there are two types of gender.
There are three types of payments.
the data is only for 2019.
we have 3 months and 31 days.
there are 6 types of product line.


```{r}
#we separate our datatypes columns we divide them into numerical and into categorical
# we had already defined categorical but need numerical

sales1_num <- subset(sales1_date, select = c(unit.price, quantity,cogs,gross.margin.percentage, gross.income, rating ,total))
head(sales1_num,n=3)
```
```{r}
#we can go ahead and change the above dataframes into tibbles for easier 
sales1_numt<-as_tibble(sales1_num)
head(sales1_numt,n=3)
```
```{R}
sales1_cat <- subset(sales1_date, select = c(branch, customer.type, gender, product.line, payment, year,month,day))
head(sales1_cat,n=3)
```


```{r}
#category columns as tible
sales1_cat_t<-as_tibble(sales1_cat)
head(sales1_cat_t,n=3)
```


# 5. Exploratory Data Analysis  (Univariate, Bivariate & Multivariate)

## 5.1 Univariate Analysis
```{r}
library(ggplot2)
library(psych)

#Descriptive analysis into measures of central Tendency

describe(sales1_numt)

```
There are 13 numerical column
the highest mean are from total and cogs variable, and lowest from quantity and rating.
the highest unit cost 100 and lowest cost 10 with most of it at 55.
the higest rating is 10 and minimum is at 4 with most of ratings at 7.
the quantiies of most unit is 5 with highest at 10 and minimum 1.


```{r}
par(mfrow = c(2, 2))
hist(sales1_numt$unit.price)
hist(sales1_numt$quantity)
hist(sales1_numt$cogs) 
hist(sales1_numt$gross.margin.percentage) 
hist(sales1_numt$gross.income) 
hist(sales1_numt$rating) 
hist(sales1_numt$total)

```

```{r}
par(mfrow = c(2, 2))
f1 <- sales1_cat$branch
f1y<- table(f1) 
head.matrix(f1y)
f2 <- sales1_cat$customer.type
f2y<- table(f2) 
head.matrix(f2y)
f3 <- sales1_cat$gender 
f3y<- table(f3)
head.matrix(f3y)
f4 <- sales1_cat$product.line 
f4y<- table(f4) 
head.matrix(f4y)
f5 <- sales1_cat$payment
f5y<- table(f5)
head.matrix(f5y)
f6 <- sales1_cat$month
f6y<- table(f6) 
head.matrix(f6y)
f7 <- sales1_cat$day 
f7y<- table(f7)
head.matrix(f7y)
```
the higest bought product is fashion accessories, and the lowest is health and beauty.
the customer types both ahve almos same number of customers.
the first month had highest sales then followed by the third and lastly the secodn month.
both genders have almost same numbers.


```{r}
# fetching the columns
products <- sales1_cat$product.line

# fetching the frequency distribution
product_frequency <- table(products)

# plotting the bargraph
barplot(product_frequency,  xlab = 'ProductRelated', ylab = 'frequency',  main = 'barplot on products')
```

Health and beauty had the lowest sales though not much worrying difference.

## 5.2 Bivariate analysis

```{r}
sales1_date %>%
ggplot() +
aes(x = day, gross.income = ..count../nrow(sales1_date), fill = gross.income) +
geom_bar() +
ylab("daily revenue trends")
```
Day 15 had most highest revenue collected followed by day 25th, day 31 had the lowest revenue collected.

```{r}
sales1_date %>%
ggplot() +
aes(x = quantity, payment = ..count../nrow(sales1_date), fill = payment) +
geom_bar() +
ylab("payment sales trends")
```


```{r}
sales1_date %>%
ggplot() +
aes(x = quantity, month = ..count../nrow(sales1_date), fill = month) +
geom_bar() +
ylab("monthly sale trends")
```
Month 1, customers who bought 10 items where high, and those who bought 8 items where the lowest.

```{r}
sales1_date %>%
ggplot() +
aes(x = quantity, product.line = ..count../nrow(sales1_date), fill = product.line) +
geom_bar() +
ylab("product sales trends")
```
most customers take 10 items of electronic accessories, fashion accessories, health & beauty and food & bevarages
6 items of home&lifestyle had most sales
7 items of sports&travel had most sales.


```{r}
sales1_date %>%
ggplot() +
aes(x = quantity, branxh = ..count../nrow(sales1_date), fill = branch) +
geom_bar() +
ylab("branch sale trends")
```
all the branches had most of its customers purches 10 items

```{r}
#Distribution of income per Branch
ggplot(sales1_date,aes(x = total, fill = branch)) +geom_density(alpha = 0.4) +labs(title = "Distribution of total sales per branch")
```
branch A and B have highest sales between 1 & 300, then all branches experience low sales toward 100 shilling.

```{r}
#Distribution of income per Gender
ggplot(sales1_date,aes(x = total, fill = gender)) +geom_density(alpha = 0.4) +labs(title = "Distribution of total sales per Gender")
```
the female buy items totaling to 250 more than males, while the males buy items totalling to a thousand more tha females.

```{r}
#Distribution of income per Gender
ggplot(sales1_date,aes(x = total, fill = customer.type)) +geom_density(alpha = 0.4) +labs(title = "does membership affect sales")
```
cystomer type-member generate more sales than the normal type.

```{r}
library(ggplot2)
library(corrplot)
```

```{R}
#Printing out correlations in our dataset
cols <-cor(sales1_num)
cols
```



```{r}
#removing the grossmargin percantage as it has std of 0.
sales1_numt[ ,c('gross.margin.percentage')] <- list(NULL)
```


```{r}
#Printing out correlations in our dataset
cols <-cor(sales1_numt)
cols
```


```{r}
corrplot(cor(cols), method="shade", tl.col="black", tl.srt=45)
```
There is an evident of positive correlation between the following columns:

cogs
gross income
total
unit price

The following columns are negatively linear:

rating
quantity


# 6. Implement the Solution

## Part 1: Dimensionality Reduction
### PCA

```{r}
#We had already ran the numericals and dropped redundant columns 
#we named it sales1_numt

pca <- prcomp(sales1_numt[,c(1:6)], center = TRUE, scale = TRUE)
summary(pca)
```
Proportion of Variance: This is the amount of variance the component accounts for in the data, ie PC1 accounts for 65.5% of total variance in the data alone!
Cumulative Proportion: This is simply the accumulated amount of explained variance, ie. if we used the first 3 components we would be able to account for 98% of total variance in the data

```{r}
#used the string function to know which variables were to be consindered
str(pca)

```
### Here we note the center point ($center), scaling ($scale),standard deviation(sdev) of each principal component. 
### The relationship (correlation or anticorrelation, etc) between the initial variables and the principal components ($rotation). 
### The values of each sample in terms of the principal components ($x)

```{r}
# We will now plot our pca. This will provide us with some very useful insights

# Installing our ggbiplot visualisation package
# 
library(devtools)
#install_github("vqv/ggbiplot")
```

```{r}
# Then Loading our ggbiplot library
#  
library(ggbiplot)
ggbiplot(pca)
```
we notice that unit price, quantity, rating and grossincome were the four variables to be highly consindered during modelling.

```{r}
#adding more details
#plotting the PCA to see the above visually
#we will use ggbiplot library
library(ggbiplot)
```

```{r}
ggbiplot(pca, obs.scale = 0.5, var.scale = 0.5, groups = sales1_date$customer.type,ellipse = TRUE, circle = TRUE)
```
we see that whether the customer is loyal or new customer the four variables will explain 65.5% plus 16.7% making it 81.2 variance which is fairly high.


## Part 2: Feature Selection

### Wrapper Method
```{r}
# Installing and loading our clustvarsel package

                         
library(clustvarsel)

```

```{r}
# Installing and loading our mclust package
library(mclust)
```

```{r}
# Sequential forward greedy search (default)
# ---
#
out = clustvarsel(sales1_numt)
out
```

we  see that the columns quantity,cogs and unitprice have been accepted and can be used for modelling.This is a more accurate way of feature selection.

```{r}
# The selection algorithm would indicate that the subset 
# we use for the clustering model is composed of variables quantity, unit price, cogs
# and that other variables should be rejected. 
# Having identified the variables that we use, we proceed to build the clustering model:
# ---
#

Subset1 = sales1_numt[,out$subset]
mod = Mclust(Subset1)
summary(mod)
```

# 7. Recommendation

From Dimentionality reduction model:
1.Quantity
2.Unit price
3.gross income
4.rating

feater selection model:
1.Quantity 
2.Cogs
3.Unit price

The feature selection model gave us the best outcome for consideration of what impact the sales in carefore kenya.

